Planck and the Extra-terrestrials\\ or what the new SI means for astronomy

### 2020-04-08 spg/main.tex
editing

\author{Prasenjit Saha} \email{psaha@physik.uzh.ch}
\affiliation{Physik-Institut, University of Zurich\\
  Winterthurerstr.~190, 8057~Zurich, Switzerland}

@@

In 1899 Planck \cite{Planck1899}.
Let $m=\sqrt{hc/G}$.
$f \rightarrow G$, $b\rightarrow h$, $a\rightarrow h/k$
Comes to $55\rm\,\mu g$.

Not the first \cite{Tomilin1999} but most prescient.  SI
\cite{Jeckelmann_2018}.  Leaves out $G$.  Also candela, sievert.

Astros use Planckian units occasionally (with $\hbar$) but generally
not.  Stories of cross-sections in gallons/Mpc.\footnote{A gallon/Mpc
  is about a kilogram, but the exact value depends on which gallon.}
Why?

@@

For astros, don't need to express solar mass in kg.  But we'll get to
that.

First Light-second: au and pc.  Again, not a new unit.  Round like Pfund.

@@

Gravity-second $GM/c^3$ (pulsars).  Strange formulations.
Dimensionless velocity ok.  But density $\rho$ as $s^{-2}$ seems
strange.  But $(G\rho)^{-1/2}$ is a time scale.  And gravity-second
per second is a power.  Basically, avoid explicit $G$ in pure
gravitational phenomena.  Explicit $G$ only when gravity and non-grav
processes both appear.

@@

\def\aj{AJ}
\def\apjl{ApJL}
\def\aap{A\&A}
\def\mnras{MNRAS}
\def\pasj{PASJ}
\def\jgr{JGR}

>>>>

\author{Prasenjit Saha} \email{psaha@physik.uzh.ch}
\affiliation{Physik-Institut, University of Zurich,
  Winterthurerstr.~190, 8057~Zurich, Switzerland}

@@

In 1899, during his struggle to understand blackbody radiation but
about a year before he found the key, Planck proposed a curious set of
units.\cite{Planck1899} In these natural units, as he called them, the
unit of mass $m=\sqrt{hc/G}$ which comes to about $55\rm\,\mu g$.  The
length unit is $h/(mc)$, the time unit is $h/(mc^2)$, while the
temperature unit is $mc^2/k$.  (Reading the original today, one needs
to substitute $f \rightarrow G$, $b\rightarrow h$, $a\rightarrow
h/k$.)  Planck then declared that these units would be recognized by
all cultures as fundamental, even extraterrestrial or non-human ones
(ausserirdische und aussermenschliche Culturen).

Planck was by no means the only person to try

Not the first \cite{Tomilin1998} but most prescient.  SI
\cite{Jeckelmann_2018}.  Leaves out $G$.  Also candela, sievert.

Astros use Planckian units occasionally\cite{magicenv} (with $\hbar$)
but generally not.\cite{dodd2011}  Why?

@@

For astronomy, the good news is that the SI no longer requires you to
express the solar mass in kg --- but we'll get to that.  We remark
first that distances in light-seconds are compliant with the new SI.
Like the electronvolt, a light-second is not itself a unit, it is an
abbreviated way of referring to the equivalent light travel time.  On
the other hand, which astronomer would wish to lose the cultural
legacies of the astronomical unit of length (au), and the parsec?
Fortunately, metrology and cultural legacies can be reconciled.  To
see how, we only to see how modern German has kept the old unit of a
Pfund, but has rounded its meaning to exactly $500\unit{g}$.  The au
and parsec have serendipitously round values: $1\unit{au}\simeq
c\times500\unit{s}$ to better than 1\%, while a parsec is just 3\%
over $c\times10^8\unit{s}$.

@@

Now we come to the kilogram, or avoiding the kilogram when necessary.
Spacecraft dynamics has long used the so-called solar mass parameter
$GM_\odot$ in $\!\unit{m}^3\unit{s}^{-2}$, which is known to eight
significant digits in Newtonian gravity, two more with general
relativity.  In pulsar timing, $GM_\odot/c^3$ in seconds is usual.  In
the new SI, by analogy with electronvolts, we can meaningfully measure
mass as $GM/c^3$ in ``gravity-seconds''.  The solar mass is about
5~micro-(gravity)-seconds.  This does not mean that kilograms are to
be avoided in astrophysics.  Rather, gravity-seconds can be useful
where kilograms would needlessly propagate the uncertainty in $G$,
which is typically the case for orbital processes.  Together with
distances in light-seconds and dimensionless velocities,
gravity-seconds allow for some elegant simplifications in the
description of all kinds of orbital processes, classical and
relativistic.\cite{2020PASP..132b1001S} Some constructions arise, that
seem {\em very} strange at first: in particular density in
gravity-seconds per cubic light-seconds, or power in gravity-seconds
per second.  But these are not really unphysical at all.  In
gravitational phenomena, a density $\rho$ is always associated with a
time scale $(G\rho)^{-1/2}$, so density as frequency squared does make
sense.  As for dimensionless power in gravity-seconds per second, this
is just power in units of the well-known Planck power $c^5/G$, which
is the luminosity scale in gravitational waves for merging black
holes, irrespective of their mass.

@@

\def\aj{AJ}
\def\apjl{ApJL}
\def\aap{A\&A}
\def\mnras{MNRAS}
\def\pasp{PASP}
